Local real space operators for translating Stokes Q/U to E/B maps on the sphere

### 2017-09-21 article/sections/numerical_implementation.tex
added new figures to 353Ghz application ; minor edits to article

In this section we present the results of evaluating the local convolution given in \eq{eq:op_qu2equbqu}, to decompose the total Stokes Q and U parameters into those correspond to the E modes and B modes of polarization respectively. We reiterate that this procedure uses real space operation on the total Stokes Q and U maps and does not go through the process of constructing the E and B modes as is usually the case. The resultant Stokes Q and U maps corresponding to the E mode are depicted in the first two rows while those corresponding to B modes are shown in the final two rows of \fig{fig:equ-bqu-maps-compare}. 

@@

\subfigure[B Q-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-qmap-healpix.pdf}}
\subfigure[B Q-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-qmap-2beta.pdf}}
\subfigure[B Q-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-qmap-diff.pdf}}
\subfigure[B U-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-umap-healpix.pdf}}
\subfigure[B U-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-umap-2beta.pdf}}
\subfigure[B U-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-umap-diff.pdf}}
\caption{The top two rows depict the Stokes Q \& U maps  respectively corresponding to E mode of polarization while the bottom two rows depict  the Stokes Q \& U maps  respectively corresponding to the  B modes of polarization. \textit{Left:} Stokes Maps derived using Healpix. \textit{Middle}: Stokes maps derived using local convolution with $r_{\rm cutoff} = 2 \beta_o$. \textit{Right:} Difference between the maps derived using the two different methods.}
\label{fig:equ-bqu-maps-compare}
\end{figure}
%
The local convolution maps depicted in this figure are computed using a radial cutoff of $r_{\rm cutoff }=2\beta_o$. Note that similar to observations of the previous section, the maps derived via local real space convolutions, suffer from having relatively higher error in the polar caps, while the differences in the  band around the equator are smaller and dominated by large scale modes.

Next we compute the E and B mode spectra from the set of Stokes maps constructed using local convolutions with different $r_{\rm cutoff}$. Note that for this exercise we use the standard Healpix routines for computing the spectra from the Stokes maps. The spectra derived from these local convolution maps and their difference with respect to the reference spectra are depicted in \fig{fig:equ-bqu-spectra_rad_cutoff}. The observations are similar to that made in the previous section: the difference with respect to the reference spectra reduce on increasing the radial cutoff and reach 2\% levels at most multipoles for the maximum radial $r_{\rm cutoff} =2 \beta_o$ used in this analysis.
%
\begin{figure}[!t] 
\centering
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/equ-spectra-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/bqu-spectra-radial-cutoff.pdf}}\\[-5ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-equ-e-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-bqu-b-spectrum-radial-cutoff.pdf}}
\caption{\textit{Top:} The thick lines in the figure on left depicts the E mode spectra derived from pure E Stokes Q and U maps derived using different radial cutoffs, while the thin lines depict the residual B mode power. The figure on the right depict the same except that the thick lines now depict the B mode spectra while the thin lines show the residual E mode power. \textit{Bottom:} The figure on the left and right depict the difference between the reference spectra and those derived from local convolution maps for different $r_{\rm cutoff}$ imposed on the convolution. In all the panels the color of the lines indicate the radial cutoff as indicated by the legend and the gray band marks the cosmic variance around the fiducial spectra.}
\label{fig:equ-bqu-spectra_rad_cutoff}
\end{figure}
%
In particular here we note that the B-mode residue in the Stokes Q \& U maps corresponding to E-modes is much smaller as compared to the E-mode residue in the Stokes Q \& U maps  computed for the B-modes of polarization. At all multipoles the B-mode residual power is below the lensing B-modes. The E-mode residual power at low multipoles is almost fully intact.

Motivated by these observations we also compute the Stokes Q \& U maps corresponding to B-modes of polarization by simply subtracting the Stokes Q \& U maps corresponding to E-mode from the total Stoker Q \& U maps. For completeness we also compute the Stokes Q \& U maps corresponding to E-mode in an analogous manner. We carry out a spectral study on these Stokes Q \& U maps corresponding to E \& B modes, computed using this alternate method. The results of the spectral analysis are summarized in \fig{fig:dbqu-spectra_rad_cutoff}.
%
\begin{figure}[!t] 
\centering
%\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/dequ-spectra-radial-cutoff.pdf}}
%\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/dbqu-spectra-radial-cutoff.pdf}} \\[-5ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-equ-de-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-bqu-db-spectrum-radial-cutoff.pdf}}
\caption{This figure depicts the same as the bottom panel of \fig{fig:equ-bqu-spectra_rad_cutoff}. The Stokes Q and U maps corresponding to E and B modes used in this analysis are constructed using the following relations: $\fqu_{E}=\fqu-\fqu_{Local}^B$, $\fqu_{B}=\fqu-\fqu_{Local}^{E}$.}
\label{fig:dbqu-spectra_rad_cutoff}
\end{figure}
%
We observe that the the low multipole spectral behavior is not affected by this alternate construction of pure Stokes Q \& U maps. The high multipole behavior is significantly altered as can be seen by comparing \fig{fig:dbqu-spectra_rad_cutoff} with lower panels of \fig{fig:equ-bqu-spectra_rad_cutoff}. The relative percentage difference with respect to the reference spectra is seen to lie well within 2\% error for sufficiently high multipoles for all radial cutoff $r_{\rm cutoff}$.
%--------------------------------------------------------
%--------------------------------------------------------
\subsection{Multipole filtering using spatial convolutions}
As we have seen in the previous section that recovering large scale modes using local convolutions is not possible. The small angular scales however are reliably recovered. However the local convolution methods can be used to recover competent spectra. 
Many CMB experiments cannot access the large angle modes due to limited sky coverage. Often the large angle analysis is deal with in pixel space owing to the non-Gaussianity of the likelihood functions. With all these as motivations we study how the radial filters are modified by removing the large scale modes. We try to understand how removing large scale modes effects the locality of the E and B map estimators. We demonstrate that it is possible to get filtered maps by using the modifying the radial filter. 
%--------------------------------------------------------
%--------------------------------------------------------

>>>>

In this section we present the results of evaluating the local convolution given in \eq{eq:op_qu2equbqu}, to decompose the total Stokes Q and U parameters into those correspond to the E modes and B modes of polarization respectively. We reiterate that this procedure uses real space operation on the total Stokes Q and U maps and does not go through the process of constructing the E and B modes as is usually the case. 

@@

\subfigure[B Q-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-qmap-healpix.pdf}}
\subfigure[B Q-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-qmap-2beta.pdf}}
\subfigure[B Q-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-qmap-diff.pdf}}
\subfigure[B U-map: Healpix]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-umap-healpix.pdf}}
\subfigure[B U-map: Local convolution]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-umap-2beta.pdf}}
\subfigure[B U-map: Difference]{\includegraphics[width=0.31\columnwidth]{simulated/qu2equbqu/b-umap-diff.pdf}}
\caption{The top two rows depict the Stokes Q \& U maps  respectively corresponding to E mode of polarization while the bottom two rows depict  the Stokes Q \& U maps  respectively corresponding to the  B modes of polarization. \textit{Left:} Stokes Maps derived using Healpix. \textit{Middle}: Stokes maps derived using local convolution with $r_{\rm cutoff} = 2 \beta_o$. \textit{Right:} Difference between the maps derived using the two different methods.}
\label{fig:equ-bqu-maps-compare}
The resultant Stokes Q and U maps corresponding to the E mode are depicted in the first two rows while those corresponding to B modes are shown in the final two rows of \fig{fig:equ-bqu-maps-compare}.  The local convolution maps depicted in this figure are computed using a radial cutoff of $r_{\rm cutoff }=2\beta_o$. Note that similar to observations of the previous section, the maps derived via local real space convolutions, suffer from having relatively higher error in the polar caps, while the differences in the  band around the equator are smaller and dominated by large scale modes.
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/equ-spectra-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/bqu-spectra-radial-cutoff.pdf}}\\[-5ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-equ-e-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-bqu-b-spectrum-radial-cutoff.pdf}}
\caption{\textit{Top:} The thick lines in the figure on left depicts the E mode spectra derived from pure E Stokes Q and U maps derived using different radial cutoffs, while the thin lines depict the residual B mode power. The figure on the right depict the same except that the thick lines now depict the B mode spectra while the thin lines show the residual E mode power. \textit{Bottom:} The figure on the left and right depict the difference between the reference spectra and those derived from local convolution maps for different $r_{\rm cutoff}$ imposed on the convolution. In all the panels the color of the lines indicate the radial cutoff as indicated by the legend and the gray band marks the cosmic variance around the fiducial spectra.}
\label{fig:equ-bqu-spectra_rad_cutoff}
Next we compute the E and B mode spectra from the set of Stokes maps constructed using local convolutions with different $r_{\rm cutoff}$. Note that for this exercise we use the standard Healpix routines for computing the spectra from the Stokes maps. The spectra derived from these local convolution maps and their difference with respect to the reference spectra are depicted in \fig{fig:equ-bqu-spectra_rad_cutoff}. The observations are similar to that made in the previous section: the difference with respect to the reference spectra reduce on increasing the radial cutoff and reach 2\% levels at most multipoles for the maximum radial $r_{\rm cutoff} =2 \beta_o$ used in this analysis.
In particular here we note that the B-mode residue in the Stokes Q \& U maps corresponding to E-modes is much smaller as compared to the E-mode residue in the Stokes Q \& U maps  computed for the B-modes of polarization. At all multipoles the B-mode residual power is below the lensing B-modes. The E-mode residual power at low multipoles is almost fully intact.

Motivated by these observations we also compute the Stokes Q \& U maps corresponding to B-modes of polarization by simply subtracting the Stokes Q \& U maps corresponding to E-mode from the total Stoker Q \& U maps. For completeness we also compute the Stokes Q \& U maps corresponding to E-mode in an analogous manner. We carry out a spectral study on these Stokes Q \& U maps corresponding to E \& B modes, computed using this alternate method. The results of the spectral analysis are summarized in \fig{fig:dbqu-spectra_rad_cutoff}.
%
\begin{figure}[!t] 
\centering
%\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/dequ-spectra-radial-cutoff.pdf}}
%\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/dbqu-spectra-radial-cutoff.pdf}} \\[-5ex]
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-equ-de-spectrum-radial-cutoff.pdf}}
\subfigure[]{\includegraphics[width=0.49\columnwidth]{simulated/qu2equbqu/relative-percentage-err-bqu-db-spectrum-radial-cutoff.pdf}}
\caption{This figure depicts the same as the bottom panel of \fig{fig:equ-bqu-spectra_rad_cutoff}. The Stokes Q and U maps corresponding to E and B modes used in this analysis are constructed using the following relations:  $[Q,U]_{E}=[Q,U]-[Q,U]_{Local}^B$ ; $[Q,U]_{B}=[Q,U]-[Q,U]_{Local}^{E}$.}
\label{fig:dbqu-spectra_rad_cutoff}
\end{figure}
%
We observe that the the low multipole spectral behavior is not affected by this alternate construction of pure Stokes Q \& U maps. The high multipole behavior is significantly altered as can be seen by comparing \fig{fig:dbqu-spectra_rad_cutoff} with lower panels of \fig{fig:equ-bqu-spectra_rad_cutoff}. The relative percentage difference with respect to the reference spectra is seen to lie well within 2\% error for sufficiently high multipoles for all radial cutoff $r_{\rm cutoff}$.
%--------------------------------------------------------
%--------------------------------------------------------
\subsection{Multipole filtering using spatial convolutions}
As we have seen in the previous section that recovering large scale modes using local convolutions is not possible. The small angular scales however are reliably recovered. However the local convolution methods can be used to recover competent spectra. 
Many CMB experiments cannot access the large angle modes due to limited sky coverage. Often the large angle analysis is deal with in pixel space owing to the non-Gaussianity of the likelihood functions. With all these as motivations we study how the radial filters are modified by removing the large scale modes. We try to understand how removing large scale modes effects the locality of the E and B map estimators. We demonstrate that it is possible to get filtered maps by using the modifying the radial filter. 
%--------------------------------------------------------
%--------------------------------------------------------
